---
title: "3 Sampling Temperature"
editor: visual
bibliography: ../references.bib
format:
  html:
    toc: true
    html-math-method: katex
embed-resources: true
---

# Sampling Temperature through Space-Time Regression Kriging

```{r}
#| label: packages
#| code-summary: "Loading packages"
#| message: false
#| warning: false
#| output: false

library(tidyverse)
packages <- c("sp", "sf", "gstat", "dbplyr", "meteo", "spacetime", "data.table", "caret",
              "terra"," ggplot2", "gganimate", "patchwork", "tictoc", "glue")
walk(packages, require, character.only = T)
```

```{r}
#| label: set_env
#| code-summary: "Set environmental variables"

set.seed(20240703) 
Sys.setenv(TZ='Europe/Stockholm')
readRenviron("../.env")
```

```{r}
load('../data/localisations/moose_random_steps.RData')
```

## 1. Retrieve Weather Data from Database

We retrieve weather stations and their data from an SMHI mock database created for this project. For details how to set up this database, view the `smhi/README.md` document. Make sure to insert credentials into the `.env` file at the repository root.

```{r}
#| label: connect to database
con <- DBI::dbConnect(
  RPostgres::Postgres(), 
  dbname = Sys.getenv("DB_NAME"),
  host = Sys.getenv("DB_HOST"),
  user = Sys.getenv("DB_USERE"),
  password = Sys.getenv("DB_PASSWORD")
)

tbl_weather_data <- tbl(con, "weather_data")
tbl_weather_stations <- tbl(con, "weather_station")
```

Let's define a function that gets us all the weather data and their weather stations between two dates, for a given parameter (e.g. air temperature). We'll later split the data into `weather_data`, and `weather_stations` separately.

```{r}
get_weather_data <- function(start_date, end_date, parameter) {
  tbl_weather_data %>% 
    filter(
      parameter == parameter,
      quality == 'G',
      date_local >= start_date,
      date_local <= end_date
    ) %>%
    left_join(tbl_weather_stations, by=c("weather_station_id"="id")) %>%
    select(
      weather_station_id, 
      date_local,
      value, 
      height, 
      geom
    ) %>%
    rename(
      station_id = weather_station_id, 
      date = date_local, 
      value = value, 
      elevation = height,
    ) %>%
    # Retrieve all records that match the criteria from the database
    collect() %>%
    mutate(
      date = as.POSIXct(date, format="%Y-%m-%d %H:%M:%S", tz="Europe/Stockholm"),
      elevation = as.integer(elevation),  # Integer is enough precision for this use case
      month = month(date),  # Month number
      year = year(date),
      day = yday(date),  # Day number from 1 - 365
      geom = st_as_sfc(geom, crs=st_crs(3006))
    ) %>%
    # Use SF to process some geospatial variables
    st_as_sf(.) %>%
    # Geometry is provided in WKT, so we need to retrieve the coords manually
    mutate(x = st_coordinates(.)[,1], y = st_coordinates(.)[,2]) %>%
    # We'll also need latitude later to calculate GTT
    st_transform(., st_crs(4326)) %>%
    mutate(lat = st_coordinates(.)[,2]) %>%
    arrange(date) %>%
    st_drop_geometry()
}
```

## 2. Regression

Space-time regression is composed of two parts. The first part consists of predicting / catching the variation of our response variable through a regression model. For our regression model we'll follow methodologies adapted from \_ and \_. We'll use elevation derived from a Swedish DEM, and geometric temperature trend (GTT). To showcase the process this notebook goes through the regression and kriging step by step.

### 2.1 Data exploration

Let's first retrieve the data.

```{r}
weather_data <- data.frame()
for (year in unique(moose_random_steps$year)) {
  start_date <- as.POSIXct(glue("{year}-06-01 00:00:00"), tz="Europe/Stockholm")
  end_date <- as.POSIXct(glue("{year}-09-01 00:00:00"), tz="Europe/Stockholm")
  yearly_weather_data <- get_weather_data(start_date, end_date, "air_temperature")
  weather_data <- bind_rows(weather_data, yearly_weather_data)
}
n_stations <- length(unique(weather_data$station_id))
weather_data$year <- as.factor(weather_data$year)  # We'll need this later
```

```{r}
# Check duplicates, generally there are none
duped <- which(duplicated(weather_data[,c("station_id", "x", "y","date")]))
print(duped)

if (length(duped) >= 1) {
  weather_data <- weather_data[-duped,]
}
```

```{r}
plot1 <- ggplot(weather_data) +
  geom_boxplot(aes(y = value, fill = factor(year)), 
               alpha = 0.8, outlier.alpha = 0.008, 
               position = position_dodge(1)) +
  scale_fill_viridis_d(option = "magma", name = "Year") +
  theme_minimal() +
  labs(y = "Temperature (°C)", title = "Boxplots of Temperatures per Year") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())

plot2 <- ggplot(weather_data %>% distinct(station_id, elevation)) +
  geom_histogram(aes(x = elevation), fill = "#FCA636FF", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Elevation", y = "Amount of Weather Stations", 
       title = glue("Histogram of Weather Station Elevation (n={n_stations})"))

combined_plot <- plot1 + plot2 + plot_layout(nrow = 2, guides = "collect")

print(combined_plot)
```

The following plot shows the temperature readings animated over a couple of days. As you can see our data is not temporally regular for all weather stations. Some weather stations update every hour, some every 3 hours.

```{r}
start_date <- min(weather_data$date)
plot_data <- weather_data %>% filter(date <= start_date + days(3))
plot_data <- st_as_sf(plot_data, coords=c('x','y'))

p <- ggplot(plot_data) + 
  geom_sf(aes(color = value, group=station_id)) +
  scale_color_viridis_c(name = "Temperature (°C)", option = "magma") +
  labs(title = "Temperature over Time",
       subtitle = "Time: {frame_time}",
       x = "Long", y = "Lat") + 
  theme_minimal() + 
  transition_time(date) +
  enter_fade() + exit_fade()

gganimate::animate(p, renderer=gifski_renderer(), duration = 25, fps=30)
```

## 2.2 Modelling

GTT is the only co variate that is also temporally dynamic, and can be calculated given a latitude and time of day. There is a handy R function called `temp_geom` from the `meteo` library that does the work for us.

```{r}
weather_data$gtt <- temp_geom(weather_data$day, weather_data$lat, variable="mean")
```

```{r}
regression_model <- lm(value ~ elevation + gtt + year, data=weather_data)
summary(regression_model)
```

As we can see, a simple linear regression is not great at predicting temperature. Nevertheless capturing this variation will prove very useful as we continue with the spatial-temporal Kriging step. For that we'll be using the residuals of this model.

Lets double-check that our residuals are normally distributed:

```{r}
plot(regression_model, which=2)
```

We see temperature tends to decrease as we increase in elevation and GTT. We'll use this linear model to fit sample and experimental semi-variograms to residuals of its prediction.

```{r}
save(regression_model, file='../data/regression_model.RData')
```

We'll also need the prediction of the model (in combination with the kriged prediction) in order to derive the temperature estimation. It's easiest to predict each moose localisation now.

```{r}
build_pred_steps <- function(data, time_col, elev_col) {
  pred <- data
  pred$day <- yday(data[[time_col]])
  pred$year <- year(data[[time_col]])
  pred$year <- factor(pred$year, levels = levels(weather_data$year))
  pred$gtt <- temp_geom(pred$day, data$lat, variable = "mean")
  pred$elevation <- data[[elev_col]]
  return(pred)
}

# Build prediction datasets for start and end steps
pred_steps_start <- build_pred_steps(moose_random_steps, "t1_", "elevation_start")
pred_steps_end <- build_pred_steps(moose_random_steps, "t2_", "elevation_end")

# Use these datasets in the regression predictions
moose_random_steps$res_start <- unname(predict(regression_model, pred_steps_start))
moose_random_steps$res_end <- unname(predict(regression_model, pred_steps_end))

# Save the updated object
save(moose_random_steps, file = "../data/localisations/moose_random_steps.RData")

rm(pred_steps_start, pred_steps_end)
```

## 3. Spatial Temporal Regression Kriging

```{r}
load('../data/regression_model.RData')
```

### 3.1 Making a test data set

Kriging using a data set of millions or rows would take a really long time. Therefore we split our data into week batches when predicting temperature at localisations. Here we'll walk through the process.

```{r}
start_date <- as.POSIXct("2016-06-01 08:00:00", tz="Europe/Stockholm")
end_date <- as.POSIXct("2016-06-08 00:00:00", tz="Europe/Stockholm")
weather_data <- get_weather_data(start_date, end_date, "air_temperature")
weather_data <- as.data.frame(weather_data)
```

### 3.1 Data preparation

We'll transform our data into a `STDF` (Space-Time Data-Frame) from the package `spacetime`. The package has a somewhat steep learning curve, but examples are provided in the [spacetime vignette](https://cran.r-project.org/web/packages/spacetime/vignettes/jss816.pdf). Another good resource is the [sftime demo](https://rdrr.io/cran/gstat/src/demo/sftime.R) of `gstat`.

```{r}
weather_data$year <- factor(weather_data$year, levels=regression_model$xlevels$year)
weather_data$gtt <- temp_geom(weather_data$day, weather_data$lat, variable="mean")
weather_data$pred <- unname(predict(regression_model, weather_data))
weather_data$res <- weather_data$value - weather_data$pred
```

A further step we could take is removing stations that are considered too close to each other. We can find the nearest station to each other station using `st_nearest_feature`. For every station that is \<100m from another, we'll remove one of the stations.

```{r}
stations <- weather_data %>% distinct(station_id, x, y)
stations_sp <- st_as_sf(stations, coords=c('x', 'y'), crs=st_crs(3006))
nearest <- st_nearest_feature(stations_sp)
nearest_df <- data.frame(
  nearest_station_id = nearest,  
  distance = as.vector(st_distance(stations_sp, stations_sp[nearest,], by_element=TRUE))
) %>% arrange(distance)

head(nearest_df)
```

We can see there's even two stations at exactly the same location, whilst we appreciate the enthusiasm this would make Kriging impossible. So instead we remove some of these stations that lie close together.

```{r}
nearest_df <- nearest_df %>% filter(distance < 100)
nearest_df <- nearest_df[duplicated(nearest_df$distance),]  # Keep one of the stations
stations_to_remove <- stations[nearest_df$nearest_station_id,]

head(stations_to_remove)

weather_data <- weather_data %>% filter(!(station_id %in% stations_to_remove$station_id))
```

We use the meteo2STFDF package to automatically enter NA for moments where stations lack a measurement For example, some stations only take measurements every 3 hours, whilst some do so every hour. This ensures that the `space-time` package sees our data as ordered, even though the intervals between stations may be different. `space-time` also supports Kriging with unordered data, but it is incredibly slow. Instead, now the NA's will instead simply be ignored by kriging. The biggest challenge is making sure variogram still fits reasonably well, and making sure there's still plenty of data even when not all stations are taking measurements.

```{r}
stations <- weather_data %>% distinct(station_id, x, y)
obs <- weather_data %>% select(station_id, date, value, elevation, gtt, res)

weather_stfdf <- meteo::meteo2STFDF(
  obs,
  stations,
  obs.staid.time = c("station_id", "date"),
  stations.staid.lon.lat = c("station_id", "x", "y"),
  crs=CRS('epsg:3006')
)
```

## 3.2 Fitting the semi-variogram

First we fit the empirical variogram, what happens under the hood is a binning of data points based on spatial and temporal difference. The squared difference in their values is then averaged over those bins, giving us the variance over spatial and temporal distance.

```{r}
# Make sure this is hours, this will be the unit for the temporal variogram
diff_time <- as.integer(diff(index(weather_stfdf@time))[1]) 
tlags = 0:ceiling(10 / diff_time)  # estimate based on visual inspection, most our data is in 1 or 3 hour intervals
max(tlags)
```

```{r}
#| output: false
emperical_variogram <- variogramST(res~1, weather_stfdf, tlags=tlags)
```

```{r}
# coerce temporal difference unit to numeric (for compatibility)
emperical_variogram$timelag <- as.numeric(emperical_variogram$timelag)
plot(emperical_variogram)
plot(emperical_variogram, wireframe=T, scales=list(arrows=F))
```

Next we attempt to fit a sum-metric space-time variogram to the empirical variogram. We base our partial sill, range, and nugget values based on visual inspection of the empirical variogram. We definitely want a small nugget effect as a lot of our weather stations are very close to each other, which could lead to the screening effect.

```{r}
linear_st_anisotropy <- estiStAni(emperical_variogram, c(100000,400000), t.range=10)

sum_metric_model <- vgmST("sumMetric",
                          space=vgm(psill=6, "Sph", range=550000, nugget=0),
                          time= vgm(psill=30, "Sph", range=10, nugget=0), 
                          joint= vgm(psill=4, "Sph",  range=50000, nugget=0),
                          stAni=linear_st_anisotropy)

fit_sum_metric_model <- fit.StVariogram(
  emperical_variogram, 
  sum_metric_model, 
  lower = c(0,
            0.001,
            0,
            0.001,  # Nugget
            1), # Range
  fit.method = 8
)
```

```{r}
# set time unit for fitted model
attr(sum_metric_model, "temporal unit") <- "hours"
# check parameters fitted model
fit_sum_metric_model
# goodness of fit sum-metric model
attr(fit_sum_metric_model, "optim.output")$value
```

```{r}
plot(emperical_variogram, fit_sum_metric_model, wireframe=T, all=T, scales=list(arrows=F))
```

### 3.3 Predicting & Cross validation

```{r}
#| output: false
#| message: false
# This file contains a function called build_weather_kriging_model which automatically
# fits a variogram model using the settings in 3.2 given some weather data. There's nothing
# new here, just the code from 3.2 written into subfunctions.
source('./03_parallel.R')
```

```{r}
# Assuming your dataframe is called `df`
folds <- createFolds(rownames(weather_data), k = 5, list = TRUE)
# Initialize an empty list to store results
cv_results <- list()
```

```{r}
head(folds$Fold1, 5)
head(folds$Fold2, 5)
```

```{r}
#| output: false
#| message: false

# Loop through each fold
for(i in 1:length(folds)) {
  
  # Create training and testing datasets
  train_indices <- unlist(folds[-i]) # use all except the current fold
  test_indices <- folds[[i]]        # the current fold for testing
  
  # Prepare the train set and model
  train_data <- weather_data[train_indices, ]
  kriging_model <- build_weather_kriging_model(train_data)
  
  # Prepare the test set
  test_data <- weather_data[test_indices, ]
  test_sf <- st_as_sf(test_data, coords = c("x", "y"), crs = CRS('epsg:3006'))
  spatial_points <- as(test_sf, "Spatial")
  test_stfdf <- stConstruct(
    test_data, 
    space = c("x", "y"), 
    time  = "date", 
    spatial_points,
    crs = CRS('epsg:3006')
  )
  
  # Kriging model for training
  tic()
  kriging_object <- krigeST(
    res~1, 
    kriging_model$weather_stfdf, 
    test_stfdf,
    kriging_model$variogram_model, 
    computeVar = T,
    maxdist = 200000,
    nmax = 10
  )
  toc()
  
  kriging_results <- kriging_object@data
  pred_combined <- kriging_results$var1.pred + kriging_results$pred
  
  # Store the results
  cv_results[[i]] <- data.frame(
    fold = i,
    true = kriging_results$value,
    pred = pred_combined,
    pred_res = kriging_results$res,
    pred_var = kriging_results$var1.var
  )
}
```

```{r}
# Combine the results into a single data frame
cv_results_df <- do.call(rbind, cv_results)
summary(cv_results_df)
```

```{r}
# Calculate R², MAE, and RMSE over all folds
cv_results_df <- cv_results_df %>%
  mutate(
    residuals = true - pred,
    squared_residuals = residuals^2,
    absolute_residuals = abs(residuals),
    total_variance = (true - mean(true))^2
  )

# Compute overall metrics (across all folds)
overall_metrics <- cv_results_df %>%
  summarise(
    mae = mean(absolute_residuals, na.rm = TRUE),
    rmse = sqrt(mean(squared_residuals, na.rm = TRUE)),
    r2 = 1 - sum(squared_residuals, na.rm = TRUE) / sum(total_variance, na.rm = TRUE)
  )

# Calculate 95% CI for all folds' predictions
ci_all_folds <- cv_results_df %>%
  summarise(
    pred_mean = mean(pred, na.rm = TRUE),
    lower = pred_mean - 1.96 * sd(pred, na.rm = TRUE),  # 95% CI lower bound
    upper = pred_mean + 1.96 * sd(pred, na.rm = TRUE),  # 95% CI upper bound
    .groups = "drop"
  )
```

```{r}
p1 <- ggplot(cv_results_df, aes(x = pred, y = true, color = as.factor(fold))) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  labs(
    x = "Predicted Values (°C)",
    y = "True Values (°C)",
    color = "Fold",
    title = "Predicted vs True (All Folds)",
    subtitle = paste(
      "R² = ", round(overall_metrics$r2, 2), " ",
      "MAE = ", round(overall_metrics$mae, 2), " ",
      "RMSE = ", round(overall_metrics$rmse, 2)
    )
  ) +
  scale_color_viridis_d(option = "magma", begin = 0.3) + 
  theme_minimal()

p1
ggsave('./figures/03_kriging_cv.jpg')
```

### 3.4 Visual inspection of results - animation over Sweden

Lets make a prediction over a time series for all of Sweden. We'll split Sweden up into a grid with a cell size of 25km. Then for each cell we'll predict the temperate at a specific time. Lastly we'll animate the values to see how temperature changes over time. First lets create the grid.

```{r, echo=F}
#| echo: false
source('03_parallel.R')
```

```{r}
#| output: false
#| message: false

# Load Sweden border polygon
sweden_border <- st_read("../data/strata/sweden_border.gpkg")
sweden_border <- st_transform(sweden_border, 3006)

# Create a regular grid within the Sweden border
grid_res <- 8000  # Grid resolution (adjust as needed)
sweden_grid <- st_make_grid(sweden_border, cellsize = grid_res)
sweden_grid <- st_intersection(sweden_grid, sweden_border)  # Keep points inside Sweden

# Convert to SpatialPointsDataFrame
sweden_grid <- terra::vect(sweden_grid)

# Load elevation raster
elevation_rast <- terra::rast("../data/geographical_covariates/elevation.tif")
elevation_at_cells <- terra::extract(elevation_rast, sweden_grid, fun=median, na.rm=T)
sweden_grid$elevation <- elevation_at_cells$elevation

# Extract coordinates
sweden_grid$x <- crds(centroids(sweden_grid))[,1]
sweden_grid$y <- crds(centroids(sweden_grid))[,2]

sweden_grid <- terra::project(sweden_grid, "epsg:4326")
sweden_grid$lat <- crds(centroids(sweden_grid))[,2]
sweden_grid$cell_id <- 1:nrow(sweden_grid)
```

Next, lets make sure these cells exists for every hour we want to Krige temperature for. You can of course make this how fine or course you want to (keeping in mind the limitations of the original data), but here we stick to one hour. For each cell-time combination we'll also calculate the covariates and the prediction of the linear regression model which are used to get the final temperate.

```{r}
sweden_df <- as.data.frame(sweden_grid)

# Define start and end times
end_date <- start_date + hours(16)
hourly_seq <- seq(start_date, end_date, by = "1 hour")
n_hours <- length(hourly_seq)

# Create a new data frame that combines the grid with hourly timestamps
sweden_rep <- sweden_df[rep(seq_len(nrow(sweden_df)), each = n_hours), ]
sweden_rep$t <- rep(hourly_seq, times = nrow(sweden_df))

# Add temporal and spatial covariates
sweden_rep$year <- factor(year(sweden_rep$t), levels = regression_model$xlevels$year)
sweden_rep$gtt <- temp_geom(yday(sweden_rep$t), sweden_rep$lat, variable = "mean")

# Predict residuals using the regression model
sweden_rep$res <- predict(regression_model, newdata = sweden_rep)

# Now sweden_grid_long is a data.frame with hourly predictions per grid cell
head(sweden_rep)
```

Lets now Krige all these cells, i'm using the original weather data as retrieved in 2.2, but you could grab a smaller or larger set of weather data here too. Please note that we're Kriging quite a lot of cells, so expect this to take a while.

```{r}
#| output: false
#| message: false

spatial_points <- st_as_sf(sweden_rep, coords=c('x', 'y'), crs=st_crs(3006))
spatial_points <- as(spatial_points, "Spatial")

# Prepare the prediction grid as a spatial-temporal object
prediction_grid_stfdf <- stConstruct(
  sweden_rep, 
  space = c("x", "y"), 
  time = 't',  # Example prediction date
  spatial_points,
  crs = CRS('epsg:3006')
)
```

```{r}
#| output: false
#| message: false

# Prepare training data using all weather stations
kriging_model <- build_weather_kriging_model(weather_data)

# Perform kriging
tic()
kriging_object <- krigeST(
  res~1, 
  kriging_model$weather_stfdf, 
  prediction_grid_stfdf,
  kriging_model$variogram_model, 
  maxdist=200000,
  nmax=10, 
  computeVar = T
)
toc()
```

Finally, lets go over our results by turning them into an animation showing how temperature changes over Sweden during our time period.

```{r}
# Convert predictions to a spatial format for visualization
kriging_results <- as.data.frame(kriging_object)
kriging_results$pred_temp <- kriging_results$res + kriging_results$var1.pred
kriging_results_sf <- st_as_sf(kriging_results, coords=c('x', 'y'), crs=st_crs(3006))

# Create the animated plot
p <- ggplot(kriging_results_sf) +
  geom_sf(aes(color = pred_temp, group = cell_id)) +
  scale_color_viridis_c(name = "Temperature (°C)", option = "magma") +
  labs(
    title = "Predicted Temperature Over Sweden",
    subtitle = "Time: {frame_time}",
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal() +
  transition_time(time) +
  enter_fade() +
  exit_fade()

# Render animation
gganimate::animate(p, renderer = gifski_renderer(), duration = 12, fps = 24)
```

## 4 Sampling weather covariates at moose steps

Next we can predict some moose steps. As we'll be doing this on large scale with multi-threading, some helper functions have been written. The `03_functions.r` file contains functions that combine all of the essentials that have been carried out so far. It also contains a small helper function to get us started with moose steps. We can either predict the temperature at the start step or the end step, here we'll show an example with end steps.he prediction of the kriging. Remember that we initially applied a power transform, so we make sure to revert that.

```{r}
#| output: false
#| message: false

prepare_end_steps <- function(df) {
  df <- df %>%
    mutate(
      x    = x2_,
      y    = y2_,
      time = t2_,
      res  = res_end
    ) %>%
    arrange(time)
  return(df)
}

# The functions shapes the data for us and adds GTT
localisations <- prepare_end_steps(
  moose_random_steps %>% 
    filter(
      t2_ > start_date, 
      t2_ < start_date + days(1)
    )
)
localisations <- as.data.frame(localisations)

# krigeST wants a ST object, lets make one
moose_sf <- st_as_sf(localisations, coords = c("x", "y"), crs = CRS('epsg:3006'))
spatial_points <- as(moose_sf, "Spatial")
moose_st <- stConstruct(
  localisations, 
  space = c("x", "y"), 
  time  = "t2_", 
  spatial_points,
  crs = CRS('epsg:3006')
)

# Start Kriging with a timer
tic()
kriging_object <- krigeST(
  res~1, 
  weather_stfdf, 
  moose_st,
  sum_metric_model, 
  maxdist=100000,
  nmax=10, 
  computeVar = T
)
toc()
```

```{r}
pred_kriging <- kriging_object@data$var1.pred
pred_combined <- pred_kriging + localisations$res_end
```

```{r}
print("Real Summary (Note: All of Sweden)")
summary(weather_data$value)
print("Reg. Kriged Summary")
summary(pred_combined)
print("Collar Temperature Summary")
summary(localisations$collar_temperature)
```

### 4.1. Using parallelisation

Below is a separate piece of code that can be used to annotate all the steps in our `moose_random_steps` object with temperature data. The run-time of this will be rather significant, so take care not to run it accidentally.

```{r, eval = FALSE}
#| eval: false

library(tidyverse)
packages <- c("sp", "sf", "gstat", "dbplyr", "meteo", "spacetime")
walk(packages, require, character.only = T)

set.seed(20240703) 
Sys.setenv(TZ='Europe/Stockholm')
readRenviron("../.env")
```

```{r, eval = FALSE}
#| eval: false
 
load('../data/localisations/moose_random_steps.RData')
load('../data/regression_model.RData')
```

```{r, eval = FALSE}
#| eval: false

# Compute the maximum number of moose data rows per batch.
max_batch_size <- 10000

# Define week boundaries.
all_times <- c(moose_random_steps$t1_, moose_random_steps$t2_)
unique_week_starts <- sort(unique(floor_date(all_times, unit = "week", week_start = 1)))
buffer_hours <- 3
unique_week_starts <- unique_week_starts[-212]

# Helper function to process one week and return a flat list of batches.
process_week <- function(week_start) {
  main_start <- week_start
  main_end   <- week_start + days(7)
  extended_start <- main_start - hours(buffer_hours)
  extended_end   <- main_end + hours(buffer_hours)
  
  # Subset moose data.
  moose_subset <- moose_random_steps %>% 
    filter(t1_ >= main_start & t1_ < main_end)
  if (nrow(moose_subset) == 0) return(NULL)
  
  # Fetch and process weather data.
  weather_data <- get_weather_data(extended_start, extended_end, "air_temperature")
  weather_data$year <- factor(weather_data$year, levels = regression_model$xlevels$year)
  weather_data$gtt <- temp_geom(weather_data$day, weather_data$lat, variable = "mean")
  weather_data$res <- weather_data$value - unname(predict(regression_model, weather_data))
  
  # Select relevant columns.
  moose_subset <- moose_subset %>% select(
    id, animal_id, step_id_, x1_, x2_, y1_, y2_, t1_, t2_, year, res_start, res_end)
  weather_data <- weather_data %>% select(station_id, date, year, x, y, value, res)
  
  # Determine the number of batches.
  num_batches <- ceiling(nrow(moose_subset) / max_batch_size)
  batches <- vector("list", num_batches)
  
  # Split into batches using a for loop.
  for (batch_num in 1:num_batches) {
    start_row <- (batch_num - 1) * max_batch_size + 1
    end_row   <- min(batch_num * max_batch_size, nrow(moose_subset))
    batches[[batch_num]] <- list(
      id = paste0(as.character(week_start), "_", batch_num),
      moose_data   = moose_subset[start_row:end_row, ],
      weather_data = weather_data
    )
  }
  return(batches)
}

# Apply the helper function to each week and flatten the result into a 1D list.
all_batches <- unlist(lapply(unique_week_starts, process_week), recursive = FALSE)

i <- 0
for (batch in all_batches){
  i <- i + nrow(batch$moose_data)
}

save(all_batches, file = '../data/all_batches.RData')
```

```{r, eval = FALSE}
#| eval: false

kriging_results <- fread('../data/kriging/03_results.csv')
moose_random_steps <- moose_random_steps %>% left_join(data, by=c("id"))
save(moose_random_steps, file='../data/localisations/moose_random_steps.RData')
```

```{r, eval = FALSE}
#| eval: false

filenames = list.files(path='../result/', pattern = '\\.csv', full.names = TRUE)
temp <- lapply(filenames, fread, sep=";")
data <- rbindlist( temp )
colnames(data) <- c("id", "temperature_start", "temperature_end")
```
