---
title: "Strata Analysis"
editor: visual
format:
  html:
    toc: true
    html-math-method: katex
embed-resources: true
bibliography: ../references.bib
---

```{r}
STRATA <- "boreal inland"
ANALYSIS_ID = "s2"
```

```{r, message=FALSE, warning=FALSE}
#| label: packages
#| code-summary: "Loading packages"
#| message: false
#| warning: false

library(tidyverse)
packages <- c("dplyr", "data.table", "amt", "terra", "sf", "sp", "ggspatial", "patchwork",
              "metR","tidyterra", "glmmTMB", "car", "survival", "mvtnorm", "glue", "ggridges",
              "tools", "scales")
walk(packages, require, character.only = T)
```

```{r}
#| code-summary: "Setting seed and loading styles"
#| message: false
#| 
set.seed(20240703) 
source('./util/styles.R')
```

## 1.0 Data Exploration

```{r}
#| code-summary: "Loading layer files"
#| message: false
#| output: false

all_strata <- st_read('../data/strata/strata.gpkg')
strata <- all_strata[all_strata$name == STRATA, ]

landuse <- rast('../data/geographical_covariates/landuse_2018.tif')
elevation <- rast('../data/geographical_covariates/elevation.tif')

landuse <- crop(landuse, strata, mask=T)
elevation <- crop(elevation, strata, mask=T)
```

```{r}
#| code-summary: "Loading moose steps"
load('../data/localisations/moose_random_steps.RData')
moose <- moose_random_steps %>% filter(strata == STRATA)
```

```{r}
# TODO remove, incomplete data due to previous bug in these years
moose <- moose %>% filter(!year %in% c(2016, 2017, 2018, 2019, 2020, 2021, 2022))
```

```{r fig.height = 5, fig.width = 8, eval=FALSE}
#| eval-summary: false
#| code-summary: "Making overview plot"

true_steps <- st_as_sf(moose[moose$case_ == 1,], coords=c("x1_", "y1_"), crs=st_crs(3006))
# Calculate the extent of the localizations
localisation_extent <- st_bbox(true_steps)  # Bounding box of the localizations
extent_sf <- st_as_sfc(localisation_extent)  # Convert bbox to an sf object
elevation <- crop(elevation, as.polygons(ext(localisation_extent), crs="EPSG:3006"), mask=T)
strata_crop <- st_crop(st_geometry(strata), localisation_extent, crs="EPSG:3006")
  
# Main plot: zoomed-in view of the localizations
localisation_elevation_plot <- ggplot(elevation, aes(x,y)) +
  # Strata
  annotation_spatial(strata_crop, fill = "transparent", color = "transparent") +  # Strata in the background
  coord_sf(xlim = c(localisation_extent["xmin"], localisation_extent["xmax"]),
           ylim = c(localisation_extent["ymin"], localisation_extent["ymax"]),
           expand = F) +
  
  # Elevation
  geom_relief(aes(z = elevation)) +
  geom_spatraster(
    data = elevation, inherit.aes = FALSE,
    aes(alpha = after_stat(value))
  ) +
  # Localizations
  layer_spatial(true_steps, alpha = 0.2, size= 0.3, color=orange) +
  
  # Scales
  scale_fill_cross_blended_c(breaks = round(seq(150, 2050, length.out=7))) +
  scale_alpha(range = c(1, 0.25)) +
  annotation_scale(width_hint=0.5, style="ticks", location="br") +
  theme_minimal() +
  theme(
    axis.title.y = element_text(vjust = +2),
    axis.title.x = element_text(vjust = -2),
    panel.grid = element_blank()
  ) + 
  guides(alpha = "none", fill = guide_legend(reverse = TRUE)) + 
  labs(
    x = "Longitude", 
    y = "Latitude", 
    fill="Elevation (m)"
  )

# Inset map: overview of strata with localization extent highlighted
strata_plot <- ggplot() +
  layer_spatial(all_strata, fill = "gray80", color = "black", alpha = 0.5) +
  layer_spatial(strata_crop, fill = orange, alpha = 0.25) +
  layer_spatial(true_steps, alpha = 0.2, size= 0.2, color=orange) +
  geom_sf(data = extent_sf, fill = NA, color = orange, linewidth = 1) +  # Highlight extent
  theme_void() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"))

combined_plot = strata_plot + 
  localisation_elevation_plot + 
  plot_layout(ncol=2) + 
    plot_annotation(
    title=glue("Localisations for {strata} strata"),
    subtitle=glue("n = {n_localisations}", n_localisations=nrow(moose))
  )

print(combined_plot)
ggsave(glue('figures/{ANALYSIS_ID}_overview.jpg'))
```

## 2. Pre-processing

### 2.1 Plot land use averaged over months

```{r, fig.height = 4, fig.width = 6}
#| code-summary: "Land cover overview for strata"
moose <- as.data.frame(moose)

# Aggregate land use proportions over an average day across multiple years
landuse_daily <- moose %>%
  mutate(avg_day = format(as.Date(t1_), "%m-%d")) %>% 
  group_by(avg_day, landuse_start) %>% 
  summarise(count = n(), .groups = "drop") %>% 
  group_by(avg_day) %>% 
  mutate(proportion = count / sum(count)) %>%  # Compute proportion per day
  ungroup()

ggplot(landuse_daily, aes(x = avg_day, y = proportion, fill = landuse_start, group = landuse_start)) +
  geom_area(alpha = 0.7) +
  scale_x_discrete(breaks = c("06-01", "07-01", "08-01", "08-31"),
                   labels = c("Jun", "Jul", "Aug", "Sep")) +  # Label months
  theme_classic() +
  labs(
    x = "Month", 
    y = "Proportion", 
    fill = "Land Cover",
    title = toTitleCase(glue("{STRATA}"))
  ) +
  landuse_legend_fill +
  guides(fill="none") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Optional: rotate labels

ggsave(glue('figures/{ANALYSIS_ID}_landcover_overview.jpg'), width=4, height=4)
```

### 2.2 Only keep dominant land uses

Only keep the most visited landuses in order to assure that each moose visited each habitat at least once.

```{r}
#| code-summary: "Pre-processing landuses"
# Identify the top 3 most frequent landuse values across both columns
top_landuses <- moose %>%
  select(landuse_start, landuse_end) %>%
  pivot_longer(cols = everything(), values_to = "landuse") %>%
  count(landuse, sort = TRUE) %>%
  slice_head(n = 3) %>%
  pull(landuse)

# Recode landuse values: keep top 3, everything else becomes "Other"
recode_top3 <- function(x, top_vals) {
  ifelse(x %in% top_vals, x, "Other")
}

# Apply the recoding
moose <- moose %>%
  mutate(
    landuse_start = factor(recode_top3(landuse_start, top_landuses),
                           levels = c(top_landuses, "Other")),
    landuse_end = factor(recode_top3(landuse_end, top_landuses),
                         levels = c(top_landuses, "Other"))
  )

unique(moose$landuse_start)
unique(moose$landuse_end)
```

```{r}
reference_landuse_category = levels(moose$landuse_end)[1]
reference_landuse_category
```

For some visual inspection, we can view each landuse visited per moose in the strata.

```{r}
moose_summary <- moose %>% 
  group_by(landuse_end, animal_id) %>% 
  summarise(tel = n(), .groups = "drop") %>% 
  arrange(animal_id)

moose_summary
```

Here we remove individuals that did not visit each habitat at least 250 times

```{r}
moose_summary <- moose_summary %>% filter(landuse_end != "Other")
ids_to_remove <- moose_summary %>% filter(moose_summary$tel < 500) %>% pull(animal_id)

print(glue("Number of moose total: ", length(unique(moose$animal_id))))
print(glue("Number of moose without 500x all habitat visits: ", length(ids_to_remove)))

moose <- moose %>% filter(!animal_id %in% ids_to_remove)
```

### 2.3 Create movement parameters

```{r}
moose <- moose %>% mutate(
  log_sl = log(sl_),
  cos_ta = cos(ta_)
)
moose$temperature_start_natural <- moose$temperature_start
moose$temperature_end_natural <- moose$temperature_end
```

### 2.4 Scale continuous variables

```{r}
cols_to_scale <- c(
  "temperature_start", "temperature_end", "distance_to_water_end",
  "elevation_end", "ruggedness_end", "skogliga_volume_end",
  "sl_", "log_sl", "cos_ta"
)

moose_unscaled <- moose %>% select(cols_to_scale)
moose_scaled <- scale(moose_unscaled)
```

```{r}
scaling_attributes <- data.frame(
  variable = cols_to_scale, 
  mean = attr(moose_scaled, "scaled:center"), 
  sd = attr(moose_scaled, "scaled:scale")
)
```

```{r}
moose <- moose %>%
  select(-all_of(cols_to_scale)) %>%  # Drop the unscaled versions
  bind_cols(moose_scaled)             # Add the scaled versions
```

### 2.5 Check for correlation between explanatory variables

```{r}
vif(
  glm(
    case_ ~ 
      sl_ +
      log_sl +
      cos_ta +
      
      temperature_end +
      distance_to_water_end +
      
      elevation_end +
      ruggedness_end +
      
      landuse_end +
      skogliga_volume_end +

      hour_s1_t2 +
      hour_s2_t2 +
      hour_s3_t2 + 
      hour_c1_t2 +
      hour_c2_t2 +
      hour_c3_t2,
      data=moose
  )
)
```

## 3 Heatwaves

### 3.1 Get heatwave events and split data in two sets

Heatwaves 3 or more days with a max temperature in the top 20th percentile

```{r}
detect_heatwave_events <- function(
    df, 
    temp_col = "temperature_start_natural", 
    time_col = "t1_",
    heatwave_percentile = 0.75, 
    min_streak_days = 3,
    control_method = c("complement", "low_percentile"),
    control_percentile = 0.2) {
  
  control_method <- match.arg(control_method)
  df <- as.data.frame(df)
  
  # Ensure 'date' column exists
  df <- df %>%
    mutate(date = as.Date(.data[[time_col]]))
  
  # Get daily maximum temperature
  daily_max <- df %>%
    group_by(date) %>%
    summarize(max_temp = max(.data[[temp_col]], na.rm = TRUE), .groups = "drop")
  
  # Determine temperature threshold for heatwave
  heat_thresh <- quantile(daily_max$max_temp, heatwave_percentile, na.rm = TRUE)
  
  # Identify candidate heatwave days
  heat_days <- daily_max %>%
    filter(max_temp > heat_thresh) %>%
    pull(date)
  
  # Helper to find consecutive streaks
  find_streaks <- function(dates, min_length) {
    dates <- sort(dates)
    streaks <- list()
    current_streak <- c()
    
    for (i in seq_along(dates)) {
      if (i == 1 || difftime(dates[i], dates[i - 1], units = "days") == 1) {
        current_streak <- c(current_streak, dates[i])
      } else {
        if (length(current_streak) >= min_length) {
          streaks <- append(streaks, list(current_streak))
        }
        current_streak <- dates[i]
      }
    }
    if (length(current_streak) >= min_length) {
      streaks <- append(streaks, list(current_streak))
    }
    return(as.Date(unlist(streaks)))
  }
  
  # Detect valid heatwave streaks
  heatwave_dates <- find_streaks(heat_days, min_streak_days)
  
  # Filter for heatwave and control data
  heatwave_data <- df %>% filter(date %in% heatwave_dates)
  
  control_data <- if (control_method == "complement") {
    df %>% filter(!date %in% heatwave_dates)
  } else {
    cold_thresh <- quantile(daily_max$max_temp, control_percentile, na.rm = TRUE)
    cold_days <- daily_max %>%
      filter(max_temp < cold_thresh) %>%
      pull(date)
    df %>% filter(date %in% cold_days)
  }
  
  return(list(
    heatwave_dates = unique(heatwave_dates),
    heatwave_data = heatwave_data,
    control_data = control_data
  ))
}

```

```{r}
result <- detect_heatwave_events(
  df = moose,
  heatwave_percentile = 0.75,
  min_streak_days = 3,
  control_method = "complement"
)
```

### 3.2 Plot Heatwaves

```{r}
plot_heatwaves <- function(
    df, 
    temp_col = "temperature_start_natural", 
    time_col = "t1_", 
    heatwave_dates,
    heatwave_percentile = 0.7) {
  
  df <- df %>%
    mutate(date = as.Date(.data[[time_col]]),
           year = year(date))
  
  # Daily max temperatures
  daily_max <- df %>%
    group_by(date, year) %>%
    summarize(max_temp = max(.data[[temp_col]], na.rm = TRUE), .groups = "drop") %>%
    arrange(year, date) %>%
    mutate(heatwave = date %in% heatwave_dates)
  
  # Group by year for proper lead calculation
  daily_max <- daily_max %>%
    group_by(year) %>%
    mutate(
      next_date = lead(date),
      next_temp = lead(max_temp),
      next_heatwave = lead(heatwave),
      segment_type = ifelse(heatwave & next_heatwave, "Heatwave", "Normal")
    ) %>%
    ungroup() %>%
    filter(!is.na(next_date))

  # Heatwave threshold
  heat_thresh <- quantile(daily_max$max_temp, heatwave_percentile, na.rm = TRUE)
  
  # Plot line segments with color based on heatwave status
  p <- ggplot(daily_max) +
    geom_segment(aes(x = date, xend = next_date,
                     y = max_temp, yend = next_temp,
                     color = segment_type),
                 linewidth = 1) +
    geom_hline(yintercept = heat_thresh, linetype = "dashed", color = "red") +
    scale_color_manual(values = c("Normal" = "grey60", "Heatwave" = "firebrick")) +
    labs(
      title = "Daily Max Temperatures with Heatwave Periods",
      subtitle = paste0("Threshold (", heatwave_percentile * 100, "th percentile): ", 
                        round(heat_thresh, 1), "°C"),
      x = "Date",
      y = "Max Daily Temperature (°C)",
      color = "Period"
    ) +
    facet_wrap(~year, scales = "free_x") +
    theme_minimal(base_size = 13)

  return(p)
}
```

```{r}
plot_heatwaves(
  df = moose,
  temp_col = "temperature_start_natural",
  time_col = "t1_",
  heatwave_dates = result$heatwave_dates,
  heatwave_percentile = 0.75
)
```

```{r}
plot_heatwaves_ridgeline <- function(
    df, 
    temp_col = "temperature_start_natural", 
    time_col = "t1_", 
    heatwave_dates,
    heatwave_percentile = 0.7) {
  
  df <- df %>%
    mutate(date = as.Date(.data[[time_col]]),
           year = factor(year(date)),
           heatwave = date %in% heatwave_dates)
  
  # Compute heatwave threshold
  heat_thresh <- quantile(df[[temp_col]], heatwave_percentile, na.rm = TRUE)
  
  # Plot: ridgeline of daily temps per year, with heatwave days filled differently
  p <- ggplot(df, aes(x = .data[[temp_col]], y = year, fill = heatwave)) +
    geom_density_ridges(scale = 2, alpha = 0.60, rel_min_height = 0.01, color = "white") +
    scale_fill_manual(values = c("TRUE" = red, "FALSE" = beige)) +
    geom_vline(xintercept = heat_thresh, linetype = "dashed", color = red) +
    labs(
      title = glue("Distributions of Daily Temperatures for {STRATA}"),
      subtitle = paste0(
        "Heatwave = >", round(heat_thresh, 1), "°C (", 
        heatwave_percentile * 100, "th percentile)"),
      x = "Daily Temperature (°C)",
      y = "Year",
      fill = "Heatwave"
    ) +
    theme_minimal(base_size = 14)

  return(p)
}

```

```{r fig.height = 8, fig.width = 7}
plot_heatwaves_ridgeline(
  df = moose,
  temp_col = "temperature_start_natural",
  time_col = "t1_",
  heatwave_dates = result$heatwave_dates,
  heatwave_percentile = 0.75
)
```

```{r}
plot_heatwave_lines <- function(
    df, 
    temp_col = "temperature_start_natural", 
    time_col = "t1_", 
    heatwave_dates,
    heatwave_percentile = 0.75
) {
  df <- df %>%
    mutate(
      date = as.Date(.data[[time_col]]),
      year = year(date),
      dummy_date = as.Date(paste0("2000-", format(date, "%m-%d")))
    ) %>%
    filter(month(date) %in% 6:8)  # Only keep summer months
  
  # Daily max temperatures
  daily_max <- df %>%
    group_by(date, dummy_date, year) %>%
    summarize(max_temp = max(.data[[temp_col]], na.rm = TRUE), .groups = "drop") %>%
    arrange(year, date) %>%
    mutate(heatwave = date %in% heatwave_dates)
  
  # Group by year for proper lead calculation
  daily_max <- daily_max %>%
    group_by(year) %>%
    mutate(
      next_dummy_date = lead(dummy_date),
      next_temp = lead(max_temp),
      next_heatwave = lead(heatwave),
      segment_type = ifelse(heatwave & next_heatwave, "Heatwave", "Normal")
    ) %>%
    ungroup() %>%
    filter(!is.na(next_dummy_date))
  
  heat_thresh <- quantile(daily_max$max_temp, heatwave_percentile, na.rm = TRUE)

  # Year-based palette for heatwave segments
  year_palette <- colorRampPalette(c("#fee0d2", "#99000d"))(length(unique(daily_max$year)))
  names(year_palette) <- sort(unique(daily_max$year))

  # Plot
  p <- ggplot() +
    # Heatwave segment highlighting
    # Daily temperature lines
    geom_line(
      data = daily_max,
      aes(x = dummy_date, y = max_temp, group = factor(year)),
      color = "grey60", linewidth = 0.6, alpha = 0.5
    ) +
    geom_segment(
      data = daily_max %>% filter(segment_type == "Heatwave"),
      aes(x = dummy_date, xend = next_dummy_date,
          y = max_temp, yend = next_temp,
          color = factor(year)),
      linewidth = 1
    ) +
    geom_hline(yintercept = heat_thresh, linetype = "dashed", color = "red") +
    scale_color_manual(values = year_palette) +
    scale_x_date(
      date_breaks = "1 month",
      date_labels = "%b",
      limits = as.Date(c("2000-06-01", "2000-08-31"))
    ) +
    labs(
      title = toTitleCase(glue("{STRATA}")),
      subtitle = paste0("Threshold (", heatwave_percentile * 100, "th percentile): ", 
                        round(heat_thresh, 1), "°C"),
      x = "Date",
      y = "Max Daily Temperature (°C)",
      color = "Heatwave Year"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      panel.grid.minor = element_blank(),
    )

  return(p)
}

```

```{r}
plot_heatwave_lines(
  df = moose,
  temp_col = "temperature_start_natural",
  time_col = "t1_",
  heatwave_dates = result$heatwave_dates,
  heatwave_percentile = 0.75
)
ggsave(glue('figures/{ANALYSIS_ID}_heatwaves_line.jpg'))
```

```{r}
moose_control <- result$control_data
moose_heatwaves <- result$heatwave_data
```

```{r}
summary(moose_control$temperature_end)
summary(moose_heatwaves$temperature_end)
```

## 3. Model fitting

### 3.1 Define model formulas with various pairs of harmonics

Simplest model, no harmonics

```{r}
model_formula_0p <- case_ ~
  landuse_end +
  distance_to_water_end +
  (landuse_end + distance_to_water_end):(temperature_end) +
  
  elevation_end + 
  ruggedness_end +
  
  sl_ +
  log_sl +
  cos_ta +
  (sl_ + log_sl + cos_ta):(temperature_start) + 
  
  strata(step_id_)
```

Model with 1 pair of harmonics

```{r}
model_formula_1p <- case_ ~  
  landuse_end + 
  distance_to_water_end + 
  elevation_end + 
  (landuse_end + distance_to_water_end + elevation_end):(hour_s1_t2 + hour_c1_t2) + 
             
  ruggedness_end + 
    
  sl_ +
  log_sl +
  cos_ta + 
  (sl_ + log_sl + cos_ta):(hour_s1_t1 + hour_c1_t1) + 
                  
  strata(step_id_)
```

Model with 2 pairs of harmonics

```{r}
model_formula_2p <- case_ ~  
  landuse_end + 
  distance_to_water_end + 
  elevation_end + 
  (landuse_end + distance_to_water_end + elevation_end):(hour_s1_t2 + hour_c1_t2 + hour_s2_t2 + hour_c2_t2) + 
             
  ruggedness_end + 
    
  sl_ +
  log_sl +
  cos_ta + 
  (sl_ + log_sl + cos_ta):(hour_s1_t1 + hour_c1_t1 + hour_s2_t1 + hour_c2_t1) + 
                  
  strata(step_id_)
```

Model with 3 pairs of harmonics

```{r}
model_formula_3p <- case_ ~  
  landuse_end + 
  distance_to_water_end + 
  elevation_end +
  (landuse_end + distance_to_water_end + elevation_end):(hour_s1_t2 + hour_c1_t2 + hour_s2_t2 + hour_c2_t2 + hour_s3_t2 + hour_c3_t2) + 
             
  ruggedness_end + 
    
  sl_ +
  log_sl +
  cos_ta + 
  (sl_ + log_sl + cos_ta):(hour_s1_t1 + hour_c1_t1 + hour_s2_t1 + hour_c2_t1 + hour_s3_t1 + hour_c3_t1) + 
                  
  strata(step_id_)
```

### 3.2 Compare model formulas (AIC), on all moose data for the strata

```{r}
model_0p <- clogit(model_formula_0p, data = moose, model = TRUE, iter.max = 40)
model_1p <- clogit(model_formula_1p, data = moose, model = TRUE, iter.max = 40)
model_2p <- clogit(model_formula_2p, data = moose, model = TRUE, iter.max = 40)
   
AIC(model_0p, model_1p, model_2p)
```

```{r}
chosen_model = model_formula_2p
```

### 3.3 Fit the best model individually for all animals

```{r}
# Function to fit conditional logistic regression models for each animal
fit_clogit_models <- function(data) {
  data %>%
    split(.$animal_id) %>%
    map(~ clogit(
      chosen_model,
      data = .x,
      model = TRUE,
      iter.max = 40
    ))
}

# Apply function to both datasets
models_control <- fit_clogit_models(moose_control)
models_heatwave <- fit_clogit_models(moose_heatwaves)
```

```{r}
get_model_summary = function(animal_id, model_list) {
  m <- model_list[[animal_id]]
  
  # Get the summary coefficients matrix.
  # The matrix rows correspond to model terms and columns to statistics.
  coefs <- summary(m)$coefficients
  coefs <- coefs[rownames(coefs) != "(Intercept)", , drop = FALSE]
  df <- data.frame(term = rownames(coefs), coefs, stringsAsFactors = FALSE)
  
  # Rename columns for clarity.
  names(df) <- c("term", "coef", "exp.coef", "std.error", "z-statistic", "p.value")
  
  # Calculate inverse variance: 1/(std.error^2)
  df$inv_var <- 1 / (df$std.error^2)
  df$id <- animal_id
  rownames(df) <- NULL
  return(df)
}

model_control_summary <- do.call(rbind, 
                                 lapply(names(models_control), 
                                        get_model_summary, 
                                        model_list = models_control)
                                 )
model_heatwave_summary <- do.call(rbind, 
                                  lapply(names(models_heatwave), 
                                         get_model_summary, 
                                         model_list = models_heatwave)
                                  )
```

```{r}
compute_population_coefs <- function(model_summary) {
  population_coefs <- model_summary %>% filter(!is.na(coef) & !is.infinite(inv_var))
  
  population_coefs <- population_coefs %>% 
    group_by(term) %>%
    group_modify(~ {
      fit = lm(coef ~ 1, data = .x %>% filter(!is.na(coef)), weights=inv_var)
      tibble(coef = coef(fit)[1])
    }) %>% 
    ungroup()
  
  return(population_coefs)
}

# Compute for both model summaries
population_coefs_control <- compute_population_coefs(model_control_summary)
population_coefs_heatwave <- compute_population_coefs(model_heatwave_summary)

# Print results
population_coefs_control
population_coefs_heatwave
```

## Bootstrap confidence intervals

```{r}
hsa_coefs <- function(model_list) {
  
  # Resample coefficients from each model
  new_coef_list <- lapply(model_list, function(l) {
    b <- coef(l)  # Point estimate
    S <- vcov(l)  # Variance-covariance matrix
    mvtnorm::rmvnorm(n = 1, mean = b, sigma = S)  # Simulated new coefficients
  })
  
  # Combine into a data frame
  new_coef <- do.call(rbind, new_coef_list)
  new_coef <- as.data.frame(new_coef) %>% 
    pivot_longer(everything(), names_to = "term", values_to = "estimate")
  
  # Fit a weighted mean model for each term
  pop_models <- new_coef %>%
    group_by(term) %>%
    filter(!any(is.na(estimate))) %>%  # Skip terms with NA values
    summarize(estimate = coef(lm(estimate ~ 1, data = cur_data()))[[1]], 
              .groups = "drop") %>%
    pivot_wider(names_from = term, values_from = estimate)

  return(pop_models)
}

get_bootstrap_coefficients <- function(model_list, population_coefs) {
    # Bootstrap with 2000 iterations for all terms
  boot <- lapply(1:2000, function(i) {
    df <- hsa_coefs(model_list)
    df$iter <- i
    return(df)
  }) %>% 
    bind_rows()
  
   # Compute bootstrap statistics
  boot_stats <- data.frame(
    lower = sapply(boot, function(x) unname(quantile(x, probs = 0.025, na.rm = TRUE))),
    mean  = sapply(boot, function(x) unname(mean(x, na.rm = TRUE))),
    upper = sapply(boot, function(x) unname(quantile(x, probs = 0.975, na.rm = TRUE)))
  )
  
  # Compute confidence interval sizes
  ci_sizes <- data.frame(
    term  = rownames(boot_stats),
    ci_lower_size = boot_stats$lower - boot_stats$mean,
    ci_upper_size = boot_stats$upper - boot_stats$mean
  )

  # Merge with population coefficients to get the final confidence intervals
  final_results <- population_coefs %>%
    rename(mean = coef) %>%  # Rename to match structure
    inner_join(ci_sizes, by = "term") %>% 
    mutate(
      lower = mean + ci_lower_size,
      upper = mean + ci_upper_size
    ) %>% 
    select(term, lower, mean, upper)  # Keep only relevant columns
  
  final_results <- final_results  %>%
    pivot_longer(cols = -term, names_to = "stat", values_to = "value") %>%
    pivot_wider(names_from = term, values_from = value) %>%
    select(-stat) %>%
    as.data.frame()
  
  rownames(final_results) <- c("lower", "mean", "upper")
  
  return(final_results)
}

boot_control_results <- get_bootstrap_coefficients(models_control, population_coefs_control)
boot_heatwave_results <- get_bootstrap_coefficients(models_heatwave, population_coefs_heatwave)
```

```{r}
compute_harmonics_scaled <- function(population_coefs) {
  # population_coefs is assumed to be a data frame where:
  # - Columns are coefficient names (e.g. "cos_ta", "hour_s1_t1", etc.)
  # - Rows are named "lower", "mean", "higher" (or stored in a column if preferred)
  
  # Create the harmonic terms (here, for t2 harmonics as in your original code)
  hour <- seq(0, 23.9, 0.1)
  hour_harmonics_df_1p <- data.frame(
    linear_term = rep(1, length(hour)),
    hour_c1_t2  = cos(2 * pi * hour / 24),
    hour_c2_t2  = cos(4 * pi * hour / 24),
    hour_s1_t2  = sin(2 * pi * hour / 24),
    hour_s2_t2  = sin(4 * pi * hour / 24)
  )
  
  # Define the groups and how to filter the coefficient names.
   groups <- list(
    "Clear-cut & young forest" = list(filter_pattern = "landuse_endClear"),
    #"Dedicious & mixed forest" = list(filter_pattern = "landuse_endDedicious"),
    "Coniferous forest" = list(filter_pattern = "landuse_endConiferous"),
    #"Open" = list(filter_pattern = "landuse_endOpen"),
    "Wetland" = list(filter_pattern = "landuse_endWetland"),
    
    "Elevation" = list(filter_pattern = "elevation"),
    "Distance to water" = list(filter_pattern = "water"),
    
    "Step length" = list(filter_pattern = "sl_", exclude_pattern = "log"),
    "Log step length" = list(filter_pattern = "log", exclude_pattern = "sl_"),
    "Cos turning angle" = list(filter_pattern = "cos")
  )
  
  # Get the statistic names from the row names of population_coefs
  stat_names <- rownames(population_coefs)
  
  # For each statistic (lower, mean, higher), compute the harmonic-scaled values
  results_list <- lapply(stat_names, function(stat) {
    df_stat <- data.frame(hour = hour)
    
    for (grp in names(groups)) {
      pat  <- groups[[grp]]$filter_pattern
      excl <- groups[[grp]]$exclude_pattern
      
      # Use the coefficient names (i.e. the column names) to select the correct terms
      coef_names <- names(population_coefs)[grepl(pat, names(population_coefs))]
      if (!is.null(excl)) {
        coef_names <- coef_names[ !grepl(excl, coef_names) ]
      }
      
      # Extract the coefficients for the given statistic and group
      # This will be a numeric vector whose length should match the number of columns in
      # hour_harmonics_df_1p
      coef_vec <- as.numeric(population_coefs[stat, coef_names, drop = FALSE])
      
      coef_mat <- matrix(coef_vec, nrow = 1)
      # Add intercept 0 to each vector if needed
      coef_vec <- apply(coef_mat, 1, function(row) {
        if (length(row) == ncol(hour_harmonics_df_1p) - 1) {
          c(0, row)  # add missing intercept
        } else {
          row
        }
      })

      # Multiply the coefficients by the harmonic terms
      # (Assumes that the order of coef_vec corresponds to the columns in hour_harmonics_df_1p)
      group_values <- as.numeric(t(as.matrix(coef_vec)) %*% t(as.matrix(hour_harmonics_df_1p)))
      
      # Add the computed values as a new column in the data frame for this statistic
      df_stat[[grp]] <- group_values
    }
    # Record which statistic these values come from
    df_stat$stat <- stat
    df_stat
  })

  # Combine the results for all statistics
  harmonics_scaled_df <- do.call(rbind, results_list)
  
  # Optionally, reshape the data into long format
  harmonics_scaled_long <- pivot_longer(
    harmonics_scaled_df,
    cols = names(groups),
    names_to = "group",
    values_to = "value"
  )
  
  return(harmonics_scaled_long)
}

harmonics_control <- compute_harmonics_scaled(boot_control_results)
harmonics_heatwave <- compute_harmonics_scaled(boot_heatwave_results)
```

## Convert to Natural Values

## Plot Results

```{r}
plot_harmonics <- function(harmonics_scaled, title, filter_groups = NULL, filter_pattern = NULL) {
  # Filter based on exact group names
  if (!is.null(filter_groups)) {
    harmonics_scaled <- harmonics_scaled %>%
      filter(group %in% filter_groups)
  }

  # Filter based on one or more regex patterns
  if (!is.null(filter_pattern)) {
    pattern_match <- Reduce(`|`, lapply(filter_pattern, function(pat) grepl(pat, harmonics_scaled$group)))
    harmonics_scaled <- harmonics_scaled[pattern_match, ]
  }

  # Convert to wide format for confidence intervals
  harmonics_wide <- harmonics_scaled %>%
    pivot_wider(names_from = stat, values_from = value)  %>%
    mutate(hour = hour+3)

  plot1 <- ggplot() +
    # Optional: uncomment if you use CI ribbons and keep group filtering consistent
    #geom_ribbon(data = harmonics_wide,
    #            aes(x = hour, ymax = upper, ymin = lower, fill = group),
    #            alpha = 0.2) +
    geom_path(data = harmonics_wide %>% select(hour, group, mean),
              aes(x = hour, y = mean, colour = group)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_x_continuous("Hour") +
    theme_classic() +
    theme(legend.position = "bottom") +
    harmonic_legend_color + 
    ggtitle(title)
  
  return(plot1)
}
```

### Plot landuse variables

```{r}
filter_pattern = levels(unique(moose$landuse_end))

p1 <- plot_harmonics(harmonics_control, "Control Model", filter_pattern = filter_pattern)
p2 <- plot_harmonics(harmonics_heatwave, "Heatwave Model", filter_pattern = filter_pattern)

p_combined <- p1 + p2 + 
  plot_annotation(
    title = toTitleCase(STRATA),
    subtitle = glue('Reference Category: {reference_landuse_category}')) + 
  plot_layout(
    ncol = 2, 
    guides = 'collect', 
    axis_titles = "collect_y") & 
  theme(legend.position = 'bottom')

p_combined <- apply_consistent_y_lims(p_combined) & 
  ylab(expression(Time-varying~parameter~values~beta))

print(p_combined)
ggsave(glue('figures/{ANALYSIS_ID}_results_habitat_selection.jpg'))
```

### Plot continuous variables

```{r}
filter_pattern = c("Elevation", "water")

p1 <- plot_harmonics(harmonics_control, "Control Model", filter_pattern = filter_pattern)
p2 <- plot_harmonics(harmonics_heatwave, "Heatwave Model", filter_pattern = filter_pattern)

p_combined <- p1 + p2 + 
  plot_annotation(title = toTitleCase(STRATA)) +
  plot_layout(
    ncol = 2, 
    guides = 'collect', 
    axis_titles = "collect_y") & 
  theme(legend.position = 'bottom')

p_combined <- apply_consistent_y_lims(p_combined) & 
  ylab(expression(Time-varying~parameter~values~beta))

print(p_combined)
ggsave(glue('figures/{ANALYSIS_ID}_results_water_elevation.jpg'))
```

### Plot movement variables

To update the Gamma and von Mises distribution from the tentative distributions (e.g. [Fieberg et al. 2021](#ref-Fieberg2021-wx), Appendix C), we just do the calculation at each time point (for the natural-scale coefficients).

Taken from notebook 1:

| strata        | gamma_shape | gamma_scale | vonmises_kappa | vonmises_mu |
|---------------|-------------|-------------|----------------|-------------|
| boreal alpine | 0.675       | 382         | 0.177          | 0           |
| boreal coast  | 0.774       | 273         | 0.0262         | 0           |
| boreal inland | 0.739       | 347         | 0.126          | 0           |
| temperate     | 0.875       | 191         | 0.0723         | 0           |

: Table with Movement Distribution Parameters per Strata

```{r}
filter_pattern = c("Step", "Log", "angle")

p1 <- plot_harmonics(harmonics_control, "Control Model", filter_pattern = filter_pattern)
p2 <- plot_harmonics(harmonics_heatwave, "Heatwave Model", filter_pattern = filter_pattern)

p_combined <- p1 + p2 + 
  plot_annotation(
    title = STRATA,
    subtitle = "Selection for movement variables over time") + 
  plot_layout(
    ncol = 2, 
    guides = 'collect', 
    axis_titles = "collect_y") & 
  theme(legend.position = 'bottom')

p_combined <- apply_consistent_y_lims(p_combined) & 
  ylab(expression(Time-varying~parameter~values~beta))

print(p_combined)
ggsave(glue('figures/{ANALYSIS_ID}_results_movement.jpg'))
```
