---
title: "04_harmonic_covariates"
format: html
editor: visual
bibliography: ../references.bib
---

```{r}
#| label: packages
#| code-summary: "Loading packages"
#| message: false
#| warning: false

library(tidyverse)
packages <- c("dplyr", "data.table", "amt", "terra", "sf", "sp", "ggspatial", "patchwork",
              "metR","tidyterra", "glmmTMB", "car", "survival", "mvtnorm")
walk(packages, require, character.only = T)
```

```{r}
#| code-summary: "Setting seed and loading styles"
#| message: false
#| warning: false
#| 
set.seed(20240703) 
source('./util/styles.R')
```

## 1.0 Data Exploration

```{r}
#| code-summary: "Loading layer files"
#| message: false
#| warning: false
STRATA <- "boreal coast"
all_strata <- st_read('../data/strata/strata.gpkg')
strata <- all_strata[all_strata$name == STRATA, ]

landuse <- rast('../data/geographical_covariates/landuse_2018.tif')
elevation <- rast('../data/geographical_covariates/elevation.tif')

landuse <- crop(landuse, strata, mask=T)
elevation <- crop(elevation, strata, mask=T)
```

```{r}
#| code-summary: "Loading moose steps"
load('../data/localisations/moose_random_steps.RData')
moose <- moose_random_steps %>% filter(strata == STRATA)
```

```{r}
# Overview of which strata plot
overview_plot <- ggplot() +
  layer_spatial(all_strata, fill = "gray80", color = "black", alpha = 0.5) +
  layer_spatial(strata, fill = "red", color = "black", size = 1) +
  labs(title = "Overview of Strata", subtitle = paste("Selected Strata:", STRATA)) +
  theme_minimal()

# Strata landuse plot
landuse_downsampled <- aggregate(landuse, fact = 50, fun = "modal", cores=7)
landuse_plot <- ggplot() + 
  layer_spatial(strata, fill = "transparent") +
  layer_spatial(as.factor(landuse_downsampled)) + 
  theme_minimal() +
  landuse_legend_fill
```

```{r}
true_steps <- st_as_sf(moose[moose$case_ == 1,], coords=c("x1_", "y1_"), crs=st_crs(3006))
# Calculate the extent of the localizations
localisation_extent <- st_bbox(true_steps)  # Bounding box of the localizations
extent_sf <- st_as_sfc(localisation_extent)  # Convert bbox to an sf object
elevation <- crop(elevation, as.polygons(ext(localisation_extent), crs="EPSG:3006"), mask=T)

# Main plot: zoomed-in view of the localizations
main_plot <- ggplot(elevation, aes(x,y)) +
  # Strata
  annotation_spatial(strata, fill = "transparent", color = black) +  # Strata in the background
  coord_sf(xlim = c(localisation_extent["xmin"], localisation_extent["xmax"]),
           ylim = c(localisation_extent["ymin"], localisation_extent["ymax"]),
           expand = F) +
  
  # Elevation
  geom_relief(aes(z = elevation)) +
  geom_spatraster(
    data = elevation, inherit.aes = FALSE,
    aes(alpha = after_stat(value))
  ) +
  # Localizations
  layer_spatial(true_steps, alpha = 0.5, color=blue) +
  
  # Scales
  scale_fill_cross_blended_c(breaks = seq(0, 400, 50)) +
  scale_alpha(range = c(1, 0.25)) +
  annotation_scale() +
  theme_minimal() +
  theme(panel.grid = element_blank()) + 
  guides(alpha = "none", fill = guide_legend(reverse = TRUE)) + 
  labs(title = "Localizations", x = "Longitude", y = "Latitude")

# Inset map: overview of strata with localization extent highlighted
inset_plot <- ggplot() +
  layer_spatial(strata, fill = "gray80", color = black) +  # All strata
  geom_sf(data = extent_sf, fill = NA, color = red, size = 5) +  # Highlight extent
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"))

# Combine the main plot and inset map
combined_plot <- main_plot +
  inset_element(
    inset_plot, 
    left = -0.75, bottom = 0.5, right = 0.99, top = 0.99, 
    align_to = "panel"
  ) 

combined_plot <- overview_plot + landuse_plot + main_plot + plot_layout(ncol=3)
print(combined_plot)
```

## 2.0 Data Selection and Scaling

For the boreal inland strata, we hypothesize that:

-   Moose select more for coniferous forest during solar noon, which provides thermal shelter

-   Moose activity is bimodal

-   Moose step length decreases as an interaction with temperature

-   Moose selection for denser forests increases with temperature

-   Moose selection for closer distances to water increases with temperature

Any start paremeters are movement kernel related, end parameters are habitat selection related

### Simple first

```{r}
moose <- moose_random_steps %>% filter(strata == STRATA)
```

```{r}
moose <- as.data.frame(moose)

# Aggregate land use proportions over an average day
landuse_daily <- moose %>%
  mutate(doy = yday(t1_)) %>%  # Extract day of year
  group_by(doy, landuse_start) %>% 
  summarise(count = n(), .groups = "drop") %>% 
  group_by(doy) %>% 
  mutate(proportion = count / sum(count))  # Compute proportion per day

# Plot the land use proportions over an average day
ggplot(landuse_daily, aes(x = doy, y = proportion, fill = landuse_start)) +
  geom_area(alpha = 0.7) +
  theme_minimal()
```

```{r}
landuse_levels = c("Forest", "Clear-cut & young forest", "Other")

moose <- moose %>%
  mutate(
    landuse_end = case_when(
      landuse_end %in% c("Clear-cut & young forest", "Open") ~ "Clear-cut & young forest",
      landuse_end %in% c("Coniferous forest", "Dedicious & mixed forest") ~ "Forest",
      TRUE ~ "Other"
    ),
    landuse_start = case_when(
      landuse_end %in% c( "Clear-cut & young forest", "Open") ~ "Clear-cut & young forest",
      landuse_start %in% c("Coniferous forest", "Dedicious & mixed forest") ~ "Forest",
      TRUE ~ "Other"
    )
  )

moose <- moose %>% mutate(
    landuse_end = factor(landuse_end, levels=landuse_levels),
    landuse_start = factor(landuse_start, levels=landuse_levels)
  )
```

```{r}
moose <- moose %>% mutate(
  log_sl = log(sl_),
  cos_ta = cos(ta_),
  
  temperature_start <- scale(temperature_start),
  temperature_end <- scale(temperature_end),
  
  skogliga_volume_end = scale(skogliga_volume_end),
  dist_to_water_end = scale(distance_to_water_end),
  
  slu_height_end = scale(slu_height_end),
  slu_volume_end = scale(slu_volume_total_end),
  slu_volume_birch_end = scale(slu_volume_birch_end),
  slu_volume_pine_end = scale(slu_volume_pine_end),
  slu_volume_spruce_end = scale(slu_volume_spruce_end),
  
  sl_ = scale(sl_),
  log_sl = scale(log_sl),
  cos_ta = scale(cos_ta)
)
```

```{r}
vif(glm(case_ ~ 
        landuse_end +
        skogliga_volume_end +
        ruggedness_end +
        temperature_end +
        dist_to_water_end + 
        sl_ +
        log_sl + 
        cos_ta + 
        hour_s1_t2 +
        hour_s2_t2 +
        hour_s3_t2 + 
        hour_c1_t2 +
        hour_c2_t2 +
        hour_c3_t2,
        data=moose))
```

### Split data into non-heatwave and heatwave

```{r}
detect_heatwave_steps <- function(df) {
  df <- as.data.frame(df)
  # Compute 95th percentile threshold
  threshold <- quantile(df$temperature_start, 0.95, na.rm = TRUE)
  # Convert t1_ to Date
  df <- df %>% mutate(date = as.Date(t1_))
  
  # Compute daily max temperature
  daily_max <- df %>% 
    group_by(date) %>% 
    summarize(max_temp = max(temperature_start, na.rm = TRUE))
  
  # Identify days above threshold
  heat_days <- daily_max %>% filter(max_temp > threshold) %>% pull(date)
  
  # Find consecutive sequences of heatwave days
  heatwave_dates <- as.Date(c())
  streak <- as.Date(c())

  if (length(heat_days) > 0) {
    for (day_index in seq_along(heat_days)) {
      if (day_index == 1) {
        streak <- c(heat_days[day_index])
        next
      }
      
      current_day <- heat_days[day_index]
      previous_day <- heat_days[day_index-1]
      if (difftime(current_day, previous_day, units = "days") == 1) {
        streak <- c(streak, current_day)
        next
      }
      if (length(streak) >= 3) {
        heatwave_dates <- c(heatwave_dates, streak)
        streak <- c(current_day)
      }
    }
  }
  
  if (length(streak) >= 3) {
    heatwave_dates <- c(heatwave_dates, streak)
  }
  
  return(heatwave_dates)
}
heatwave_dates <- detect_heatwave_steps(moose)
moose_control <- moose %>% filter(!as.Date(t1_) %in% heatwave_dates)
moose_heatwaves <- moose %>% filter(as.Date(t1_) %in% heatwave_dates)
```

### Fit models for individual animals

```{r}
model_formula_simple <- case_ ~
  landuse_end +
  dist_to_water_end +
  (landuse_end + dist_to_water_end):(temperature_end) +
  
  ruggedness_end +
  
  sl_ +
  log_sl +
  cos_ta +
  (sl_ + log_sl + cos_ta):(temperature_end) + 
  
  strata(step_id_)
```

```{r}
model_formula_1p <- case_ ~  
  landuse_end + 
  dist_to_water_end + 
  (landuse_end + dist_to_water_end):(hour_s1_t2 + hour_c1_t2) + 
             
  # MOVEMENT KERNEL
  ruggedness_end + 
    
  sl_ +
  log_sl +
  cos_ta + 
  (sl_ + log_sl + cos_ta):(hour_s1_t1 + hour_c1_t1) + 
                  
  strata(step_id_)
```

```{r}
model_formula_2p <- case_ ~  
  landuse_end + 
  dist_to_water_end + 
  (landuse_end + dist_to_water_end):(hour_s1_t2 + hour_c1_t2 + hour_s2_t2 + hour_c2_t2) + 
             
  # MOVEMENT KERNEL
  ruggedness_end + 
    
  sl_ +
  log_sl +
  cos_ta + 
  (sl_ + log_sl + cos_ta):(hour_s1_t1 + hour_c1_t1 + hour_s2_t1 + hour_c2_t1) + 
                  
  strata(step_id_)
```

```{r}
model_formula_3p <- case_ ~  
  landuse_end + 
  dist_to_water_end + 
  (landuse_end + dist_to_water_end):(hour_s1_t2 + hour_c1_t2 + hour_s2_t2 + hour_c2_t2 + hour_s3_t2 + hour_c3_t2) + 
             
  # MOVEMENT KERNEL
  ruggedness_end + 
    
  sl_ +
  log_sl +
  cos_ta + 
  (sl_ + log_sl + cos_ta):(hour_s1_t1 + hour_c1_t1 + hour_s2_t1 + hour_c2_t1 + hour_s3_t1 + hour_c3_t1) + 
                  
  strata(step_id_)
```

## Compare model formulas (AIC)

```{r}
model_simple <- clogit(model_formula_simple, data = moose, model = TRUE, iter.max = 40)
model_1p <- clogit(model_formula_1p, data = moose, model = TRUE, iter.max = 40)
model_2p <- clogit(model_formula_2p, data = moose, model = TRUE, iter.max = 40)
model_3p <- clogit(model_formula_3p, data = moose, model = TRUE, iter.max = 40)
   
AIC(model_simple, model_1p, model_2p, model_3p)
```

## Fit models for individuals

```{r}
models_control <- moose_control %>% 
  split(.$animal_id) %>% 
  map(function(individual_data) {
    clogit(
      model_formula_3p,
      data = individual_data,
      model = TRUE,
      iter.max = 40
    )
  })

models_heatwave <- moose_heatwaves %>% 
  split(.$animal_id) %>% 
  map(function(individual_data) {
    clogit(
      model_formula_3p,
      data = individual_data,
      model = TRUE,
      iter.max = 40
    )
  })
```

```{r}
get_model_summary = function(animal_id, model_list) {
  m <- model_list[[animal_id]]
  
  # Get the summary coefficients matrix.
  # The matrix rows correspond to model terms and columns to statistics.
  coefs <- summary(m)$coefficients
  coefs <- coefs[rownames(coefs) != "(Intercept)", , drop = FALSE]
  df <- data.frame(term = rownames(coefs), coefs, stringsAsFactors = FALSE)
  
  # Rename columns for clarity.
  names(df) <- c("term", "coef", "exp.coef", "std.error", "z-statistic", "p.value")
  
  # Calculate inverse variance: 1/(std.error^2)
  df$inv_var <- 1 / (df$std.error^2)
  df$id <- animal_id
  rownames(df) <- NULL
  return(df)
}

model_control_summary <- do.call(
  rbind, lapply(names(models_control), get_model_summary, model_list = models_control))
model_heatwave_summary <- do.call(
  rbind, lapply(names(models_heatwave), get_model_summary, model_list = models_heatwave))
```

```{r}
compute_population_coefs <- function(model_summary) {
  population_coefs <- model_summary %>% filter(!is.na(coef) & !is.infinite(inv_var))
  
  population_coefs <- population_coefs %>% 
    group_by(term) %>%
    group_modify(~ {
      fit = lm(coef ~ 1, data = .x %>% filter(!is.na(coef)), weights=inv_var)
      tibble(coef = coef(fit)[1])
    }) %>% 
    ungroup()
  
  return(population_coefs)
}

# Compute for both model summaries
population_coefs_control <- compute_population_coefs(model_control_summary)
population_coefs_heatwave <- compute_population_coefs(model_heatwave_summary)

# Print results
print(population_coefs_control)
print(population_coefs_heatwave)

```

## Bootstrap confidence intervals

```{r}
hsa_coefs <- function(model_list) {
  
  # Resample coefficients from each model
  new_coef_list <- lapply(model_list, function(l) {
    b <- coef(l)  # Point estimate
    S <- vcov(l)  # Variance-covariance matrix
    mvtnorm::rmvnorm(n = 1, mean = b, sigma = S)  # Simulated new coefficients
  })
  
  # Combine into a data frame
  new_coef <- do.call(rbind, new_coef_list)
  new_coef <- as.data.frame(new_coef) %>% 
    pivot_longer(everything(), names_to = "term", values_to = "estimate")
  
  # Fit a weighted mean model for each term
  pop_models <- new_coef %>%
    group_by(term) %>%
    filter(!any(is.na(estimate))) %>%  # Skip terms with NA values
    summarize(estimate = coef(lm(estimate ~ 1, data = cur_data()))[[1]], 
              .groups = "drop") %>%
    pivot_wider(names_from = term, values_from = estimate)

  return(pop_models)
}

get_bootstrap_coefficients <- function(model_list, population_coefs) {
    # Bootstrap with 2000 iterations for all terms
  boot <- lapply(1:2000, function(i) {
    df <- hsa_coefs(model_list)
    df$iter <- i
    return(df)
  }) %>% 
    bind_rows()
  
   # Compute bootstrap statistics
  boot_stats <- data.frame(
    lower = sapply(boot, function(x) unname(quantile(x, probs = 0.025, na.rm = TRUE))),
    mean  = sapply(boot, function(x) unname(mean(x, na.rm = TRUE))),
    upper = sapply(boot, function(x) unname(quantile(x, probs = 0.975, na.rm = TRUE)))
  )
  
  # Compute confidence interval sizes
  ci_sizes <- data.frame(
    term  = rownames(boot_stats),
    ci_lower_size = boot_stats$lower - boot_stats$mean,
    ci_upper_size = boot_stats$upper - boot_stats$mean
  )

  # Merge with population coefficients to get the final confidence intervals
  final_results <- population_coefs %>%
    rename(mean = coef) %>%  # Rename to match structure
    inner_join(ci_sizes, by = "term") %>% 
    mutate(
      lower = mean + ci_lower_size,
      upper = mean + ci_upper_size
    ) %>% 
    select(term, lower, mean, upper)  # Keep only relevant columns
  
  final_results <- final_results  %>%
    pivot_longer(cols = -term, names_to = "stat", values_to = "value") %>%
    pivot_wider(names_from = term, values_from = value) %>%
    select(-stat) %>%
    as.data.frame()
  
  rownames(final_results) <- c("lower", "mean", "upper")
  
  return(final_results)
}

boot_control_results <- get_bootstrap_coefficients(models_control, population_coefs_control)
boot_heatwave_results <- get_bootstrap_coefficients(models_heatwave, population_coefs_heatwave)
```

```{r}
compute_harmonics_scaled <- function(population_coefs) {
  # population_coefs is assumed to be a data frame where:
  # - Columns are coefficient names (e.g. "cos_ta", "hour_s1_t1", etc.)
  # - Rows are named "lower", "mean", "higher" (or stored in a column if preferred)
  
  # Create the harmonic terms (here, for t2 harmonics as in your original code)
  hour <- seq(0, 23.9, 0.1)
  hour_harmonics_df_1p <- data.frame(
    linear_term = rep(1, length(hour)),
    hour_c1_t2  = cos(2 * pi * hour / 24),
    hour_c2_t2  = cos(4 * pi * hour / 24),
    hour_c3_t2  = cos(6 * pi * hour / 24),
    hour_s1_t2  = sin(2 * pi * hour / 24),
    hour_s2_t2  = sin(4 * pi * hour / 24),
    hour_s3_t2  = sin(6 * pi * hour / 24)
  )
  
  # Define the groups and how to filter the coefficient names.
  # Adjust these patterns so they match the names of the coefficients in your bootstrap results.
  groups <- list(
    "Dist.Water" = list(filter_pattern = "water"),
    "sl_" = list(filter_pattern = "sl_", exclude_pattern = "log"),
    "log(sl_)" = list(filter_pattern = "log", exclude_pattern = "sl_"),
    "cos(ta_)" = list(filter_pattern = "cos")
  )
  
  # Get the statistic names from the row names of population_coefs
  stat_names <- rownames(population_coefs)
  
  # For each statistic (lower, mean, higher), compute the harmonic-scaled values
  results_list <- lapply(stat_names, function(stat) {
    df_stat <- data.frame(hour = hour)
    
    for (grp in names(groups)) {
      pat  <- groups[[grp]]$filter_pattern
      excl <- groups[[grp]]$exclude_pattern
      
      # Use the coefficient names (i.e. the column names) to select the correct terms
      coef_names <- names(population_coefs)[grepl(pat, names(population_coefs))]
      if (!is.null(excl)) {
        coef_names <- coef_names[ !grepl(excl, coef_names) ]
      }
      # Extract the coefficients for the given statistic and group
      # This will be a numeric vector whose length should match the number of columns in
      # hour_harmonics_df_1p
      coef_vec <- as.numeric(population_coefs[stat, coef_names, drop = FALSE])
      # Multiply the coefficients by the harmonic terms
      # (Assumes that the order of coef_vec corresponds to the columns in hour_harmonics_df_1p)
      group_values <- as.numeric(t(as.matrix(coef_vec)) %*% t(as.matrix(hour_harmonics_df_1p)))
      
      # Add the computed values as a new column in the data frame for this statistic
      df_stat[[grp]] <- group_values
    }
    # Record which statistic these values come from
    df_stat$stat <- stat
    df_stat
  })
  
  # Combine the results for all statistics
  harmonics_scaled_df <- do.call(rbind, results_list)
  
  # Optionally, reshape the data into long format
  harmonics_scaled_long <- pivot_longer(
    harmonics_scaled_df,
    cols = names(groups),
    names_to = "group",
    values_to = "value"
  )
  
  return(harmonics_scaled_long)
}

harmonics_control <- compute_harmonics_scaled(boot_control_results)
harmonics_heatwave <- compute_harmonics_scaled(boot_heatwave_results)
```

```{r}
plot_harmonics <- function(harmonics_scaled, title) {
  # Convert to wide format for confidence intervals
  harmonics_wide <- harmonics_scaled %>%
    pivot_wider(names_from = stat, values_from = value)
  
  ggplot() +
    geom_ribbon(data = harmonics_wide %>% filter(!group %in% c("Clear-cut & young Forest")), 
                aes(x = hour, ymax=upper, ymin=lower, fill = group), 
                alpha = 0.2) +
    geom_path(data = harmonics_wide %>% select(hour, group, mean), 
              aes(x = hour, y = mean, colour = group)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_y_continuous(expression(Time-varying~parameter~values~beta)) +
    scale_x_continuous("Hour") +
    theme_classic() +
    theme(legend.position = "bottom") +
    ggtitle(title)
}

# Generate individual plots
p1 <- plot_harmonics(harmonics_control, "Control Model")
p2 <- plot_harmonics(harmonics_heatwave, "Heatwave Model")

p_combined <- p1 + p2 + plot_layout(ncol=2)

# Print the final combined plot
print(p_combined)
```

```{r}
library(mgcv)
moose_heatwaves$times <- 1
fit_slopes <- gam(cbind(times, step_id_) ~ 
                    landuse_end + 
                    s(hour_t2, by = landuse_end, bs = "cc", k=5) + 
                    
                    distance_to_water_end +
                    s(hour_t2, by = distance_to_water_end, bs = "cc", k=5) + 
                    
                    sl_ + 
                    log_sl + 
                    cos_ta +
                    s(hour_t1, by = sl_, bs = "cc", k=5) +
                    s(hour_t1, by = cos_ta, bs = "cc", k=5),
                  data = moose_heatwaves,
                  family = cox.ph, 
                  weights = y)
```

```{r}
summary(fit_slopes)
```

```{r}
gratia::draw(fit_slopes, rug = FALSE) 
```

# Old Code

## Landuse

```{r}
{r}
moose %>% 
  count(landuse_start, name = "count") %>% 
  arrange(desc(count))

# Generate all possible combinations of animal IDs and land use types
all_combinations <- expand.grid(
  animal_id = unique(moose$animal_id), 
  landuse_start = unique(moose$landuse_start)
)

# Identify land use types that were never visited for each animal
visited <- moose %>% distinct(animal_id, landuse_start)

missing_landuse <- anti_join(
    all_combinations, visited, 
    by = c("animal_id", "landuse_start")
  ) %>% 
  pull(landuse_start) %>% 
  unique()

cat("Missing land use types:\n", paste(missing_landuse, collapse = ", "), "\n\n")

# Define land use types and animals to filter out
unused_landuse <- c("Open", "Water bodies", "Anthropogenic", "Wetland")

# Apply filters
moose <- moose %>% 
  filter(
    !landuse_start %in% unused_landuse,
    !landuse_end %in% unused_landuse
  )
moose <- as.data.frame(moose)
# Count occurrences of each land use type per animal
moose_summary <- moose %>% 
  group_by(landuse_end, animal_id) %>% 
  summarise(tel = n(), .groups = "drop")
```

## Bootstrapping

```{r}
{r}
# Load packages
library(amt)
library(dplyr)
library(tidyr) # to pivot longer
library(mvtnorm) # for multivariate normal random number generator

# Load sample data
dat <- amt_fisher
hab <- amt_fisher_covar

amt_fisher_covar$elevation <- terra::unwrap(amt_fisher_covar$elevation)
amt_fisher_covar$popden <- terra::unwrap(amt_fisher_covar$popden)

# Fit individual models
mods <- dat %>% 
  nest(trk = x_:t_) %>% 
  mutate(hsf = map(trk, function(xx) {
    xx %>% 
      random_points() %>% 
      extract_covariates(amt_fisher_covar$elevation) %>% 
      extract_covariates(amt_fisher_covar$popden) %>% 
      # Give available points large weights when fitting HSF (not iSSF)
      mutate(weight = case_when(
        case_ ~ 1,
        !case_ ~ 1e5
      )) %>% 
      glm(formula = case_ ~ elevation + popden, data = ., family = binomial,
          weights = weight)
  }))

covs <- mods %>%
  mutate(coef = map(hsf, tidy)) %>% 
 select(sex, id, name, coef) %>% 
  unnest(cols = coef) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(inv_var = 1/(std.error^2))

elev <- lm(estimate ~ 1, data = covs, subset = term == "elevation",
           weights = inv_var)
summary(elev)

# Model population-level popden coefficient
# Intercept-only model = population (weighted) mean
popden <- lm(estimate ~ 1, data = covs, subset = term == "popden",
           weights = inv_var)
summary(popden)


## Basically the same as above until here; now we write a function we can use
## when bootstrapping. There are probably countless ways to go about this
## that would all give identical results. This is one example.

# Function to return population-level coefficients for popden and elev
# We will pass it the list of individual models so we can access the
# coefficients and variance-covariance matrix from each model object.

hsa_coefs <- function(model_list) {
  
  # Resample the individual coefficients from each model
  new_coef_list <- lapply(model_list, function(l) {
    # Point estimate
    b <- coef(l)
    # Variance-covariance matrix
    S <- vcov(l)
    # Return new coefficients
    b_new <- mvtnorm::rmvnorm(n = 1, mean = b, sigma = S)
  })
  
  # Combine in a data.frame
  new_coef <- as.data.frame(do.call(rbind, new_coef_list)) %>% 
    pivot_longer(everything(), names_to = "term", values_to = "estimate")
  
  # Model population-level elevation coefficient
  # Dropping the weights because bootstrapping is propagating the uncertainty
  # Intercept-only model = population mean
  # Would be equivalent to use mean(), but this allows addition of covariates
  elev <- lm(estimate ~ 1, data = new_coef, subset = term == "elevation")
  
  # Model population-level popden coefficient
  # Dropping the weights because bootstrapping is propagating the uncertainty
  # Intercept-only model = population mean
  # Would be equivalent to use mean(), but this allows addition of covariates
  popden <- lm(estimate ~ 1, data = new_coef, subset = term == "popden")
  
  # Return coefficients in data.frame
  return(data.frame(elev = coef(elev)[[1]], popden = coef(popden)[[1]]))
}

# (If you wanted to, you could calculate log-RSS inside that function)

# Now bootstrap to get many sets of new coefficients.
# Using 100 iterations for speed, but you probably want 2000 + for inference
set.seed(1)
boot <- lapply(1:100, function(i) {
  df <- hsa_coefs(mods$hsf)
  df$iter <- i
  return(df)
}) %>% 
  bind_rows()

upr_boot <- quantile(boot$`landuse_endClear-cut & young forest`, 0.975)
lwr_boot <- quantile(boot$`landuse_endClear-cut & young forest`, 0.025)

# Now that we're done sampling, we can calculate log-RSS by hand
x1 <- data.frame(elev = 120, popden = 1000)
x2 <- data.frame(elev = 80, popden = 500)

# Still calculate point estimate for log-RSS using original model fit
# Linear predictors (sometimes called g(x))
g1 <- coef(elev)[[1]] * x1$elev + coef(popden)[[1]] * x1$popden
g2 <- coef(elev)[[1]] * x2$elev + coef(popden)[[1]] * x2$popden

# log-RSS
log_rss <- g1 - g2

# Now get confidence interval from bootstrap sample

# Linear predictors for bootstrap iterations
#   Note that this short code snippet only works with nrow(x1) == 1
#   Otherwise, iterate over rows of x1.
boot$g1 <- boot$elev * x1$elev + boot$popden * x1$popden
boot$g2 <- boot$elev * x2$elev + boot$popden * x2$popden

# log-RSS for bootstrap iterations
boot$log_rss <- boot$g1 - boot$g2

# Mean of bootstrap sample
mean_boot <- mean(boot$log_rss)

# For 95% confidence interval:
# 97.5th quantile
upr_boot <- quantile(boot$log_rss, 0.975)
# 2.5th quantile
lwr_boot <- quantile(boot$log_rss, 0.025)

# Distance from bootstrap mean to lwr/upr bounds gives *size* of the CI
# (Note that mean_boot != log_rss)
upr_dist <- upr_boot - mean_boot
lwr_dist <- lwr_boot - mean_boot

# Construct confidence interval by adding distances to MLE ('log_rss')
c(log_rss + lwr_dist, 
  est = log_rss,
  log_rss + upr_dist)

```
